---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Goal of this analysis:

-   Step 1) Cluster taxa in trait databases according to their trait profiles to obtain groups of taxa with similar trait profiles (= trait profile groups)

-   Step 2) Extract those traits that drive the clustering/are responsible for the grouping of the taxa in groups of similar trait profiles

#### Two Analyses:

-   Analysis A) Comparison between Europe, North America, Australia, and New Zealand.

    -   Aggregation to family level required because of varying taxonomic resolutions between trait databases

    -   Subset of grouping features/traits that are present in all databases (feeding mode, locomotion, respiration, body size, body form, reproduction/oviposition, voltinism)

-   Analysis B) Comparison between major climatic regions of Europe and North America

    -   Occurrence data for Europe based on Ecoregions

    -   Occurrence data for North America from recently published CONUS trait database (Genus occurrences)

    -   Aggregation of trait information to Genus level required

### Data analysis and current challenges

-   Data analysis:

    -   For Step 1) Hierarchical cluster analysis:

        -   Traits are expressed as affinities based on fuzzy coding

        -   Normalised to a range between $[0-1]$

        -   Distance matrix based on **Orloci's Chord distance** (= Euclidean distance after scaling vector length to 1)

            -   Solves problems of:

                -   Species having different total sums of trait affinities

                -   Joint absences

#### Small hypothetical example Orloci's Chord distance

```{r}
library(vegan)
library(dplyr)

# Example data
ex_trait <- data.frame(
  feed_shredder = c(0, NA, 0.7, 0),
  feed_gatherer = c(0.1, NA, 0.3, 0),
  feed_predator = c(0, NA, 0, 0),
  feed_parasite = c(0, NA, 0, 0),
  feed_filter = c(0, NA, 0, 0),
  feed_herbivore = c(0.9, NA, 0, 1),
  resp_teg = c(NA, 0.5, 0, 0),
  resp_gil = c(NA, 0.5, 0, 0.5 ), 
  
  resp_pls_spi = c(NA, 0, 1, 0.5), 
  
  volt_semi = c(NA, 0, 0, 0),
  volt_uni = c(NA, 0, 1, 1),
  volt_bi_multi = c(NA, 1, 0, 0))

# Orloci's chord distance
dist_mat <- decostand(ex_trait, "norm", na.rm = TRUE) %>%
  vegdist(., "euclidean", na.rm = TRUE) %>%
  as.matrix()

# Standardized to range [0, 1]
dist_mat/sqrt(2)

# convert NAs to zero
ex_trait_zero <- vapply(ex_trait, function(y)
  ifelse(is.na(y), 0, y), FUN.VALUE = c(
    "1" = 0,
    "2" = 0,
    "3" = 0,
    "4" = 0
  ))

# With zeros instead of NAs
dist_mat_zero <- decostand(ex_trait_zero, "norm") %>%
  vegdist(., "euclidean") %>%
  as.matrix()

dist_mat_zero/sqrt(2)
```

#### Data availability

-   European taxa:

```{r}

# TODO only show output but hide code
library(ggplot2)
library(data.table)

#### Functions used --------------------------------------------------------------------------------
create_pattern_ind <- function(x, non_trait_cols) {
  if (missing(non_trait_cols)) {
    trait_names_pattern <- sub("\\_.*|\\..*", "", names(x)) %>%
      unique() %>%
      paste0("^", .)
  } else{
    pat <- paste0(non_trait_cols, collapse = "|")
    # get trait names & create pattern for subset
    trait_names_pattern <-
      grep(pat, names(x), value = TRUE, invert = TRUE) %>%
      sub("\\_.*|\\..*", "", .) %>%
      unique() %>%
      paste0("^", .)
  }
  trait_names_pattern
}

#### check for completeness of trait dataset ####
completeness_trait_data <- function(x, non_trait_cols) {
  trait_names_pattern <- create_pattern_ind(
    x = x,
    non_trait_cols = non_trait_cols
  )

  # test how complete trait sets are
  output <- matrix(ncol = 2, nrow = length(trait_names_pattern))
  for (i in seq_along(trait_names_pattern)) {
    # vector containing either 0 (no NAs) or a number (> 0) meaning that all
    # entries for this trait contained NA
    vec <-
      x[, apply(.SD, 1, function(y) {
        base::sum(is.na(y))
      }),
      .SDcols = names(x) %like% trait_names_pattern[[i]]
      ]

    # How complete is the dataset for each individual trait?
    output[i, ] <-
      c(
        (length(vec[vec == 0]) / nrow(x)) %>%
          `*`(100) %>%
          round(),
        trait_names_pattern[[i]]
      )
  }
  return(as.data.frame(output))
}

#### Load European trait data ----------------------------------------------------------------------
# aggregated to genus level
# classified according to different climatic regions
trait_eu <- readRDS(
  file.path(
    getwd(),
    "Comparison_between_climatic_zones",
    "Cache",
    "preproc_data_genus.rds"
  )
)

# Data availability
trait_coverage <- lapply(
  trait_eu,
  FUN = function(y) {
    completeness_trait_data(
      x = as.data.table(y)
    )
  }
)


trait_coverage <- rbindlist(trait_coverage, idcol = "Climateregion") 
setnames(trait_coverage,
  old = c("V1", "V2"),
  new = c("Coverage", "Grouping_feature")
)
trait_coverage[, Coverage := as.numeric(Coverage)]

# plot
gf_descend <- trait_coverage[
  order(Climateregion, -Coverage),
  unique(Grouping_feature)
]
trait_coverage[, Grouping_feature := factor(Grouping_feature,
  levels = gf_descend
)]

ggplot(trait_coverage) +
  geom_pointrange(aes(
    x = as.factor(Grouping_feature),
    y = Coverage,
    ymin = 0,
    ymax = Coverage
  )) +
  facet_wrap(~Climateregion) +
  labs(x = "Grouping feature", 
       y = "% Taxa that contain trait information") +
  scale_x_discrete(
    labels = c(
      "Feeding mode",
      "Voltinism",
      "Respiration",
      "Locomotion",
      "Oviposition",
      "Size",
      "Body form"
    )
  ) +
  coord_flip() +
  theme_bw() +
  theme(
    legend.position = "none",
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11)
  )
```

-   **Q1**: Should I not consider entries with NAs, leave them as they are, or substitute with Zeros? 
If I leave them as they are taxa the distance measure will be calculated based on the amount of trait information available.
If I substitute NAs with Zeros, this will change the distances when calculating the distance matrix (see example taxon 1 and 4). Replacing NAs with zeros seems to be wrong from an ecological standpoint, because this indicates that a particular taxon would not express certain trait/traits.






