---
title: "Tables convergence trait profiles - Overview trait data"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, 
                    output_dir = "/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Paper/Tables/") })
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE}
# Set up
library(data.table)
library(dplyr)
library(reactable)
library(htmltools)
```

# Continental trait data

## Table 2
Overview of families of aggregated continental trait data. Shown are the total number of families per trait dataset and the number of families per order.


```{r, echo = FALSE, message = FALSE}
trait_data_bind <-
  readRDS(
    "/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Cache/trait_data_ww_bind.rds"
  )

# Calculate nr of taxa & prop of orders
trait_data_bind[, nr_taxa := .N, by = "continent"]
trait_data_bind[, nr_fam_per_order := .N, by = .(continent, order)]
trait_data_bind[, prop_order := nr_fam_per_order / nr_taxa]

cont_td_summary <-
  unique(trait_data_bind[, .(continent, nr_taxa, order, nr_fam_per_order)]) %>%
  dcast(., ... ~ order)

reactable(
  cont_td_summary,
  columns = list(
    continent = colDef(name = "Continent",
                       width = 110),
    nr_taxa = colDef(name = "Total nr. of families"),
    Ephemeroptera = colDef(width = 150)
  ),
  highlight = TRUE
)
```

# Trait data climate regions

## Table 3

Overview of genera of aggregated European climatic region trait dataset.

```{r, echo = FALSE, message = FALSE}
# EU
eu_clim <-
  readRDS(
    "/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Cache/Cache_comparison_climatic_reg/eu_glvl_murria_pp.rds"
  )

eu_clim_summary <-
  eu_clim[, .(genus, family, order, continent, climateregion)]
eu_clim_summary[, nr_genera_cr := .N, by = "climateregion"]
eu_clim_summary[, nr_genera_cr_order := .N, by = .(climateregion, order)]
eu_clim_summary <-
  eu_clim_summary[, .(order, family, nr_genera_cr_order, nr_genera_cr, climateregion)] %>%
  dcast(.,
        order + nr_genera_cr_order + nr_genera_cr + climateregion ~ family,
        fun.aggregate = length) %>%
  .[order(climateregion),]

reactable(
  eu_clim_summary,
  columns = list(
    order = colDef(name = "Order",
                   width = 100),
    nr_genera_cr_order = colDef(name = "Nr. of genera per climateregion and order"),
    nr_genera_cr = colDef(name = "Total nr. genera per climateregion"),
    climateregion = colDef(name = "Climateregion")
  ),
  filterable = TRUE,
  highlight = TRUE
)

```

### Table 4:

Overview of genera of aggregated North American climatic region trait dataset.

```{r, echo = FALSE, message = FALSE}
# NOA
noa_clim <-
  readRDS(
    "/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Cache/Cache_comparison_climatic_reg/noa_glvl_pp.rds"
  )

noa_clim_summary <- noa_clim[, .(genus, family, order, continent, climateregion)]
noa_clim_summary[, nr_genera_cr := .N, by = "climateregion"]
noa_clim_summary[, nr_genera_cr_order := .N,
                 by = .(climateregion, order)]
noa_clim_summary <-
  noa_clim_summary[, .(order, family, nr_genera_cr_order, nr_genera_cr, climateregion)] %>%
  dcast(.,
        order + nr_genera_cr_order + nr_genera_cr + climateregion ~ family,
        fun.aggregate = length) %>%
  .[order(climateregion),]

reactable(
  noa_clim_summary,
  columns = list(
    order = colDef(name = "Order",
                   width = 120),
    nr_genera_cr_order = colDef(name = "Nr. of genera per climateregion and order", 
                                width = 120),
    nr_genera_cr = colDef(name = "Total nr. genera per climateregion"),
    climateregion = colDef(name = "Climateregion")
  ),
  filterable = TRUE,
  highlight = TRUE
)
```