---
title: "Tables convergence trait profiles"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Paper/Tables/") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
# Set up
library(data.table)
library(reactable)
library(htmltools)
```

## Table 2

Contribution of grouping features to the global distances.

```{r, echo = FALSE}
# Load data global distances
global_dist <- readRDS("/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Cache/global_dist.rds")
global_dist <- rbindlist(global_dist, idcol = "continent")

reactable(global_dist[, .(continent, grouping_feature, global_dist = round(`global distance`, digits = 2))],
          columns = list(continent = colDef(name = "Continent"), 
                         grouping_feature = colDef(name = "Grouping feature"),
                         global_dist = colDef(name = "Global distance",
                                              align = "center")), 
          filterable = TRUE,
          highlight = TRUE)
```

## Table 3

Five most important traits to distinguish TPGs for each continent. After the initial run the highest ranking variable has been removed and a random forest has been fitted again. This procedure has been repeated 4 times. Training error and test error for the fitted random forest are shown in the last two columns.

```{r, echo = FALSE}

rf_vimp <- readRDS("/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Cache/rf_vimp.rds")
reactable(
  rf_vimp[, .(
    continent,
    five_most_important_traits,
    run,
    test_error = round(test_error, digits = 2),
    train_error = round(train_error, digits = 2)
  )],
  columns = list(
    continent = colDef(name = "Continent",
                       width = 120),
    five_most_important_traits = colDef(name = "Five most important traits",
                                        width = 300),
    run = colDef(name = "Run",
                 width = 250),
    test_error = colDef(name = "Test error"),
    train_error = colDef(name = "Train error")
  ), 
  filterable = TRUE,
  highlight = TRUE
)
```

## Table 4

Brier score for test and training data of the random forest models.

```{r}

perf_summary <- readRDS("/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Cache/perf_summary.rds")

```

## Table 5

```{r, echo = FALSE}

summary_cr_taxa <- readRDS("/home/kunzst/Dokumente/Projects/Trait_DB/Convergence-trait-profiles/Cache/Cache_comparison_climatic_reg/summary_cr_taxa.rds")

reactable(
  summary_cr_taxa,
  highlight = TRUE, 
  columns = list(
    `Nr. Taxa EU` = colDef(name = "Nr. Taxa EU (mostly species lvl)"),
    `Nr. Taxa NOA` = colDef(name = "Nr. Taxa NOA (genus lvl)")
  ))
```
